---
title: "R Markdown for Novices"
author: "Heike W"
date: 2018-11-03T21:13:14-05:00
categories: ["R"]
tags: ["R Markdown", "R"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

I strongly encourage every researcher and every analyst to use it, because R Markdown will make your life easier:


- Easily share your result: R Markdown can convert your results to various formats such as html, word and pdf. No more copy pasting and everything is up to date!

- Quickly explore your data: You use R Markdown like a Notebook and see can see multiple Outputs and plots without having to use the RStudio Viewer back buttons.

- Ensure your work is replicable: You can save your results (e.g. a plot or a table) and the code that created it, in one file

- Automate reports: Analyses that need to be in the consistent and in the same format everytime can be automated.


### So how does it work?
In an R Markdown File (.RMD) you combine R code chunks and markdown text. Everything is knitted together to a Markdown format, which can be converted in almost any format through Pandoc. The outputfile may contain your code, descriptions (executed Markdown) and the R output (tables, plots, numbers...). 

-- insert figure

There are 3 different components of R Markdown files:

#### YAML Header
Your Header contains meta data about your .RMD File, such as the author, date and output format. The header is on top of every .RMD file. The code is enclosed with 3 dashes. 
```{r yaml, eval = FALSE}
--- 
title: "My first R Markdown document"
author: "Heike W"
date: 2018-11-03T21:13:14-05:00
output: html_document
---
```


#### Markdown (Text)
Markdown is a simple textformatting language. Similar to HTML or Latex, you write plain text. Punctuation characters indicate the formating.

# Header #

#### R Code Chunks
The third component, your R Code. You can write any code, you would normally write in R in R markdown. R code is enclosed with 3 apostrophies, followed by curly braces. In the curly braces, you can tell R, what it should do with the code?

|How should the output look like?| How to change the parameter| Is it the default option?|
|---------------------------------------|--------------------------|--------------------|
|include code | echo = TRUE | yes|
|automatically format code| tidy = TRUE | no|
|do not execute code | eval = FALSE | no|
|execute code, but show nothing | include = FALSE| no|
|do not show any warnings| warings = FALSE | no|
|do not show any errors | error = FALSE | no|

```{r}
myData <- iris
head(iris)
```



