---
title: Using R Markdown for recurring Reports
author: Heike
date: '2018-11-05'
slug: using-r-markdown-for-recurring-reports
categories: []
tags: ["R","R Markdown", "Reporting","Analytics"]
---

In this post I will show you some of my best practises for using R Markdown for reporting. We will cover

1. How to create beautiful tables with conditional formating with kableExtra

2. How to include your results in the (Markdown) test

3. How to include global variables.

# Create beautiful tables with kableExtra

`kableExtra` is a nice library that allows you to format and style your tables. It workds similar to `ggplot2`: You create a base table and then add formating layers with the pipe operator `%>%`.

````r
`r ''````{r myTable}
library(kableExtra)
head(iris) %>%
  kable() %>%
  kable_styling()
```
````

```{r, echo = FALSE}
library(kableExtra)
head(iris) %>%
  kable() %>%
  kable_styling()
```



By default, tables take 100 % of the page width.If you have only a few columns, this looks weird. You create less wide tables by setting ` full_width = F`

````r
`r ''````{r myTable}
library(kableExtra)
head(iris[1:2]) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```
````

```{r, echo = FALSE}
library(kableExtra)
head(iris[1:2])  %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

You can pass various arguments to `kable_styling` to influence the Font and the position of the table.

|Format | Description | Example | 
|--------------------------|-------------------------------------|-------------------------------|
|**Font**|||
|Font size| change size of font| font_size = 7|
|Font| ..| ..|
|**Position**|||
|Position table| Align the table left ("left"), right ("right"), center ("center")|`position = "left"`|
|Embed table in text| wrap text around the table| `position = "float_right"`|

Let's increase the font size and position the table on the left.

````r
`r ''````{r myTable}
library(kableExtra)
head(iris[1:2]) %>%
  kable() %>%
  kable_styling(font_size = 20, position = "left", full_width = FALSE)
```
````

```{r, echo = FALSE}
library(kableExtra)
head(iris[1:2]) %>%
  kable() %>%
  kable_styling(font_size = 20, position = "left", full_width = FALSE)
```
